Key Derivation From Noisy Sources With More Errors Than Entropy

### 2014-09-17 trunk/constantRate/ConstCompFuzzExt-reorg.tex
intro fixing

Constructions of fuzzy extractors are limited by the tension between security and correctness guarantees: if we allow for higher error tolerance $t$, then we also need higher starting entropy $m$. The reason for this tension is simple: if an adversary who knows $p$ can guess any $w'$ within distance $t$ of $w$, then it will be able to easily obtain the true $r$ by running $\rep$.
%If $t$ is larger, then a single guess for $w'$ may be within distance $t$ of more $w$ values, thus increasing adversarial probability of winning. 
In fact, if $t$ is high enough that there are $2^m$ points in a ball of radius $t$, then there exists a distribution of $w$ of min-entropy $m$  \emph{contained entirely in a single ball}.  For this distribution, an adversary can run $\rep$ on the center of this ball and always learn the key $r$.
Thus, if the security guarantee of a given fuzzy extractor holds for \emph{any} source of a given min-entropy $m$ and the correctness guarantees holds for any $t$ errors, then $m$ must   be greater than $\log |B_t|$, where $|B_t|$ denotes the number of points in a ball of radius $t$.  If a source fails this condition, we will says that it has \emph{more errors than entropy}.

@@

%Prime candidate sources for authentication have $\Huse\le 0$.  
Unfortunately, prime candidate sources for authentication have more errors than entropy. 
As an example, the iris is believed to be the best biometric for high security applications~\cite{prabhakar2003biometric}.  Daugman~\cite{daugman2004} designs a transform called IrisCode (using specialized wavelets) to derive a $2048$ bit string from an iris.  Let the outcome of this transform (on different irises) define a distribution $w$.  The precise number of errors that must be tolerated depends on the desired correctness.  For correctness of around $20\%$, a $t$ of approximately $205$ is required. Thus,
\[
\log |B_t|
= \log \sum_{i=0}^{205} {2048 \choose i} \approx 956\,.
\]
In contrast, $w$ is estimated to have about $249$ bits of entropy~\cite{daugman2004}.
There is considerable subsequent research~\cite{gentile2009slic,gentile2009efficient,rathgeb2011combining}, but it does not affect the above comparison dramatically.%\footnote{The work of Hao et al.~\cite[Section 4.3]{hao2006combining} provides a similar analysis but their calculation underestimates the number of possible error patterns; their calculation, which we cannot confirm, is $\Huse \approx 44$.}

The situation with reusability is even worse: the only known construction of reusable fuzzy extractors \cite{Boyen2004} requires very particular relationships between $w_i$ values, which are unlikely to hold in any practical source. \bnote{this should be strengthened.  move negative results to there.}

@@

%Fuller, Meng, and Reyzin show that replacing secure sketches with a similar computational component is unlikely to be fruitful~\cite[Corollary 3.8, Theorem 3.10]{fuller2013computational}.  Instead, they suggest two alternatives: produce a new consistent secret with computational entropy instead of recovering $w$~(our first construction) or combine the information-reconciliation and privacy amplification components~(our second construction).

Most known constructions of fuzzy extractors put sufficient information in $p$ to recover the original $w$ from a nearby $w'$ during $\rep$ (this procedure is called a \emph{secure sketch}), and then apply a randomness extractor to $w$ to get $r$.
As shown by~\cite{DBLP:journals/siamcomp/DodisORS08}\lnote{specific theorem number} for the information-theoretic case and by \cite[Corollary 3.8, Theorem 3.10]{fuller2013computational} for the computational case, this is unlikely to work for sources with more errors than entropy. Thus, we give up on recovering $w$. \bnote{I'm not sure I buy this argument.  We have coding bounds, but they don't say that secure sketches won't work.}

Our first construction reduces the alphabet size by hashing each input symbol (which comes from a large alphabet) into a much smaller set, so that the resulting hash value has low entropy deficiency.
The intuition behind this approach is that it reduces the size of $B_t$ by reducing the alphabet size, but preserves a sufficient portion of the input entropy.  The resulting string no longer has more errors than entropy.
 We then apply any information-theoretic fuzzy extractor to the resulting string

@@

Information-theoretic fuzzy extractors are obtained by replacing computational distance by statistical distance.  We do make a second definitional modification.  The standard definition of information-theoretic fuzzy extractors considers $\mathcal{W}$  consisting of all distributions of a given entropy.  As described in the introduction, it is impossible to provide security for all distributions with more errors than entropy.  Thus, in both the computational and information-theoretic settings we consider arbitrary families of distributions.

@@

Ideally, a fuzzy extractor allows a user to authenticate to multiple different services/devices.  That is, instead of creating a single pair $(r, p)$ from an initial reading $w$, we read the noisy source multiple times~($w_1,..., w_q$) and produce $(r_1, p_1),..., (r_q, p_q)$.  This problem is known as a reusable fuzzy extractor~\cite{Boyen2004}.  Boyen considers two versions of the problem, first where the adversary sees $p_1,..., p_q$~(outsider security).  Second where the adversary controls some subset of the servers and sees key generation on arbitrary $p$~(insider security).  
Boyen constructs a outsider reusable cryptographic fuzzy extractor based on the code-offset or the syndrome sketches from~\cite{JW99, DBLP:journals/siamcomp/DodisORS08} instantiated with linear codes. Boyen also transforms this construction in an insider secure construction using random oracles. 
\paragraph{Previous Negative Results}

There are two primary questions about the reusability of a fuzzy extractor.  The first is how often it can be reused, the second is what correlation is allowed between $w_1,..., w_q$.  Boyen~\cite{Boyen2004} considers fuzzy extractors that are reusable an unbounded number of times.  Their construction is secure when the correlations between $w_1,.., w_q$ come from a subgroup of permutations used in the construction of the fuzzy extractor.  They then show information-theoretic fuzzy extractors secure with unbounded reuse can be transformed into their construction~\cite[Theorem 11]{Boyen2004}.  This shows that correlation between $w_1,..., w_q$ must be restricted to this limited class.  There is no reason to expect that noise between repeated readings comes from a well-behaved subgroup of permutations.  Their result can be interpreted as saying that reusability is impractical.

However, this result is inherently information-theoretic, the adversary must exhaust all randomness of the fuzzy extractor and computes statistical distance over exponential size strings.  There are two  restrictions to avoid this result: 1) limit the number of reuses of the fuzzy extractor 2) provide security against computational adversaries.  The second restriction implies the adversary will only observe a polynomial number of invocations of the fuzzy extractor.

We limit the adversary to be computationally bounded~(and thus sees only a limited number of reuses).
%We present a non-adaptive definition and note that our results extend to the adaptive definition presented by Boyen~\cite[Definition 7]{Boyen2004}.\bnote{We need to come back to justifying this or provide a proof that our scheme works in the adaptive setting.}
%\begin{definition}[Insider Reusable Fuzzy Extractors]
%\label{def:reusable fuzz}
%Let $\mathcal{W}$ be a family of distributions over $\mathcal{M}$.  Let $(\gen, \rep)$ be a $(\mathcal{M}, \mathcal{W}, \kappa, t)$-computational fuzzy extractor that is $(\epsilon, s)$-hard with error $\delta$.  Let $W_1,..., W_q$ be a joint distribution such that $\forall i, W_i\in\mathcal{W}$.  Let $(R_i, P_i)$ be the distribution formed by running $\gen$ on the distribution $W_i$.  We say that $(\gen, \rep)$ is $q$-reusable if for all $i$
%\begin{align*}
%\delta^{\mathcal{D}_{s}}(&(R_1, P_1), ..., (R_i, P_i),..., (R_q, P_q), \\
%&(R_1, P_1),..., (U_\kappa, P_i),..., (R_q, P_q)) < \epsilon.
%\end{align*}
%\end{definition}
Our definition is a modified form of the definition of Boyen~\cite[Definition 7]{Boyen2004}.  We discuss definitional differences in \apref{sec:comparing reusability}.

>>>>

Constructions of fuzzy extractors are limited by the tension between security and correctness guarantees: if we allow for higher error tolerance $t$, then we also need higher starting entropy $m$. The reason for this tension is simple: if an adversary who knows $p$ can guess any $w'$ within distance $t$ of $w$, then it will be able to easily obtain the true $r$ by running $\rep$.
%If $t$ is larger, then a single guess for $w'$ may be within distance $t$ of more $w$ values, thus increasing adversarial probability of winning. 
In fact, if $t$ is high enough that there are $2^m$ points in a ball of radius $t$, then there exists a distribution of $w$ of min-entropy $m$  \emph{contained entirely in a single ball}.  For this distribution, an adversary can run $\rep$ on the center of this ball and always learn the key $r$.
Thus, if the security guarantee of a given fuzzy extractor holds for \emph{any} source of a given min-entropy $m$ and the correctness guarantees holds for any $t$ errors, then $m$ must   be greater than $\log |B_t|$, where $|B_t|$ denotes the number of points in a ball of radius $t$.  This condition on the source is necessary regardless of whether the fuzzy extractor is information-theoretic or computational, and extends even to the interactive setting.
If a source fails this condition, we will says that it has \emph{more errors than entropy}.

@@

%Prime candidate sources for authentication have $\Huse\le 0$.  
Unfortunately, prime candidate sources for authentication have more errors than entropy. 
For example, the IrisCode~\cite{daugman2004}, which is the state of the art approach to handling what is believed to be the best biometric \cite{prabhakar2003biometric}, produces a source that more errors than entropy~\cite[Section 5]{blanton2009biometric}. Physically inclonable functions suffer from similar problems \lnote{ben please fix}

%As an example, the iris is believed to be the best biometric for high security applications~%\cite{prabhakar2003biometric}.  Daugman~\cite{daugman2004} designs a transform called IrisCode (using specialized wavelets) to derive a $2048$ bit string from an iris.  Let the outcome of this transform (on different irises) define a distribution $w$.  The precise number of errors that must be tolerated depends on the desired correctness.  For correctness of around $20\%$, a $t$ of approximately $205$ is required. Thus,
%\[
%\log |B_t|
%= \log \sum_{i=0}^{205} {2048 \choose i} \approx 956\,.
%\]
%In contrast, $w$ is estimated to have about $249$ bits of entropy~\cite{daugman2004}.
%There is considerable subsequent research~\cite{gentile2009slic,gentile2009efficient,rathgeb2011combining}, but it does not affect the above comparison dramatically.%\footnote{The work of Hao et al.~\cite[Section 4.3]{hao2006combining} provides a similar analysis but their calculation underestimates the number of possible error patterns; their calculation, which we cannot confirm, is $\Huse \approx 44$.}

The situation with reusability is even worse: the only known construction of reusable fuzzy extractors \cite{Boyen2004} requires very particular relationships between $w_i$ values, which are unlikely to hold in any practical source. \bnote{this should be strengthened.  move negative results to there.} \lnote{waiting for response form Adam}

@@

%Fuller, Meng, and Reyzin show that replacing secure sketches with a similar computational component is unlikely to be fruitful~\cite[Corollary 3.8, Theorem 3.10].  Instead, they suggest two alternatives: produce a new consistent secret with computational entropy instead of recovering $w$~(our first construction) or combine the information-reconciliation and privacy amplification components~(our second construction).

Most known constructions of fuzzy extractors put sufficient information in $p$ to recover the original $w$ from a nearby $w'$ during $\rep$ (this procedure is called a \emph{secure sketch}), and then apply a randomness extractor to $w$ to get $r$.
Unfortunately, the current techniques for building secure sketches do not work for sources with more errors than entropy, because they lose at least $\log |B_t|$ bits of entropy regardless of the source (this loss is necessary when the source is uniform~\cite[Lemma C.1]{DBLP:journals/siamcomp/DodisORS08}). Moreover, relaxing the security definition to computational will not improve this approach~\cite[Corollary 3.8, Theorem 3.10]{fuller2013computational}. Thus, we take a different approach and do not attempt to recover $w$.

Our first construction reduces the alphabet size by hashing each input symbol (which comes from a large alphabet) into a much smaller set, so that the resulting hash value has low entropy deficiency.
The intuition behind this approach is that it reduces the size of $B_t$ by reducing the alphabet size, but preserves a sufficient portion of the input entropy.  The resulting string no longer has more errors than entropy.
 We then apply any information-theoretic fuzzy extractor to the resulting string.

@@

Information-theoretic fuzzy extractors are obtained by replacing computational distance by statistical distance.  We do make a second definitional modification.  The standard definition of information-theoretic fuzzy extractors considers $\mathcal{W}$  consisting of all distributions of a given entropy.  As described in the introduction, it is impossible to provide security for all distributions with more errors than entropy.  Thus, in both the computational and information-theoretic settings we consider limited families of distributions.

@@

Ideally, a fuzzy extractor allows a user to authenticate to multiple different services/devices.  That is, instead of creating a single pair $(r, p)$ from an initial reading $w$, we read the noisy source multiple times~($w_1,..., w_q$) and produce $(r_1, p_1),..., (r_q, p_q)$.  This problem is known as a reusable fuzzy extractor~\cite{Boyen2004}.  There are two primary parameters of the definition: 1) $q$, the reusability of the fuzzy extractor and 2) what class of correlation is allowed between readings $w_i$ and $w_j$.

Boyen considers two versions of the problem, first where the adversary sees $p_1,..., p_q$~(outsider security).  Second where the adversary controls some subset of the servers and sees key generation on arbitrary $p$~(insider security).  
Boyen constructs a outsider reusable cryptographic fuzzy extractor for unbounded $q$ when the correlation between readings is limited to a transitive isometric permutation groups. Boyen transforms this construction to insider security using random oracles. In both constructions, the correlation between  $w_1,..., w_q$ is limited to a transitive isometric permutation group.  Unfortunately, this class of correlation is relatively weak and unrealistic.
%\paragraph{Previous Negative Results}

Boyen~\cite{Boyen2004} considers information-theoretic fuzzy extractors with unbounded reusability.  
However, even for this relatively weak class of correlations, Boyen shows that reusability comes at a real cost~\cite[Theorem 11]{Boyen2004}.  If the correlation between readings contains a set that generates a transitive isometric permutation group, any secure sketch must lose entropy proportional to the number of errors corrected.\footnote{This result strengthens the bounds of~\cite[Appendix C]{DBLP:journals/siamcomp/DodisORS08}, who show this loss is necessary for uniform $W_1$.} This means achieving reusability for sources with more errors than entropy is impossible in this setting.
%\bnote{old stuff}
%Their construction is secure when the correlations between $w_1,.., w_q$ come from a subgroup of permutations used in the construction of the fuzzy extractor.  They then show information-theoretic fuzzy extractors secure with unbounded reuse can be transformed into their construction~\cite[Theorem 11]{Boyen2004}.  This shows that correlation between $w_1,..., w_q$ must be restricted to this limited class.  There is no reason to expect that noise between repeated readings comes from a well-behaved subgroup of permutations.  Their result can be interpreted as saying that reusability is impractical.
%
%However, this result is inherently information-theoretic, the adversary must exhaust all randomness of the fuzzy extractor and computes statistical distance over exponential size strings.  There are two  restrictions to avoid this result: 1) limit the number of reuses of the fuzzy extractor 2) provide security against computational adversaries.  The second restriction implies the adversary will only observe a polynomial number of invocations of the fuzzy extractor.

To overcome this result, we consider computationally bounded adversaries.
%We present a non-adaptive definition and note that our results extend to the adaptive definition presented by Boyen~\cite[Definition 7]{Boyen2004}.\bnote{We need to come back to justifying this or provide a proof that our scheme works in the adaptive setting.}
%\begin{definition}[Insider Reusable Fuzzy Extractors]
%\label{def:reusable fuzz}
%Let $\mathcal{W}$ be a family of distributions over $\mathcal{M}$.  Let $(\gen, \rep)$ be a $(\mathcal{M}, \mathcal{W}, \kappa, t)$-computational fuzzy extractor that is $(\epsilon, s)$-hard with error $\delta$.  Let $W_1,..., W_q$ be a joint distribution such that $\forall i, W_i\in\mathcal{W}$.  Let $(R_i, P_i)$ be the distribution formed by running $\gen$ on the distribution $W_i$.  We say that $(\gen, \rep)$ is $q$-reusable if for all $i$
%\begin{align*}
%\delta^{\mathcal{D}_{s}}(&(R_1, P_1), ..., (R_i, P_i),..., (R_q, P_q), \\
%&(R_1, P_1),..., (U_\kappa, P_i),..., (R_q, P_q)) < \epsilon.
%\end{align*}
%\end{definition}
We modify the definition of Boyen~\cite[Definition 7]{Boyen2004}.  We discuss the differences in \apref{sec:comparing reusability}.
